---
title: "Stats and Figures"
author: "Vanessa Clark"
date: "29/07/2020"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

#Libraries
```{r include=FALSE}
#Statistics
library(brms)           #Bayesian regression models 
library(broom)          #Takes messy outputs of built-in R function band turns them into tidy data frames
library(car)            #"Companion to applied regresion"
library(dplyr)          #For working with data frame like objects
library(emmeans)        #To do calculations for matrix, use after fitting models to explore models more and for predictions
library(ggeffects)      #Compute marginal effects from statistical models and returns the result as tidy data frames, "ggemmeans", "ggeffect"
library(lme4)           #Also for mixed effects models
library(lubridate)      #Simplification working with dates and times
library(MASS)           #"stepAIC" - Modern Applied Statistics with S
library(mgcv)           #Mixed GAM (generalised additive models)
library(mnormt)         #Multivariate, normal and t distributions
library(multcomp)       #Test and confidence intervals for general linear hypotheses and parametric models
library(MuMIn)          #"r.squaredGLMM" Tools for performing model selection and model averaging
library(nlme)           #for linear and non-linear mixed effects models
library(psych)          #Basic descriptive stats + multivariate and scale construction using factor analysis
library(rlang)          #A toolbox for working with base types, core R features like the condition system, and core 'Tidyverse' features like tidy evaluation.
library(Rmisc)          #Data Analysis and utility operations
library(sjlabelled)     #Working with labelled data - can move from SPSS, statisica etc. 
library(sjmisc)         #Data Transformation functions
library(sjstats)        #Linear models, or linear mixed effects models not in base r
library(survival)       #Survival analysis, Surv objects, Kaplan-Meier, and Alan-johansen curves
library(tidybayes)      #tidy data and geoms for bayesion models
library(tidyr)          #Data tidying ("spread", and "gather" primary functions)


#Graphics
library(cowplot)        #Creating figures "plot_grid"
library(DHARMa)         #Create simulation based residuals for fitted linear mixed models "plotResiduals 
library(effects)        #Effect displays for linear and generalized linear, and other models
library(egg)            #Arranging multiple plots
library(ggfortify)      #Data Visualization Tools for Statistical Analysis Results
library(ggplot2)        #Creating data visualisation
library(ggpubr)         #ggarrange, "Publication ready plots" based off ggplot figures
library(grid)           #Graphics layout
library(gridExtra)      #To work with "grid" graphics, to arrange multiple grid-based plots on a page, and draw tables
library(jtools)         #Tools for plotting regression analysis, works with "survey" package
library(lattice)        #Data visualisation
library(RColorBrewer)   #colour palettes for graphics
library(scales)         #Help scales and other functions like maps and others
library(sjPlot)         #data visualisation
library(survminer)      #survival curves using ggplot2
library(tidyverse)      #Collection of packages such as ggplot, dplyr, tidyr
library(vcd)            #visualising categorical data


```
#Functions

##Model diagnostics
```{r}

#Combined model diagnostics

Model.Diagnostics <- function(model1, anovatype=2){
      S.Results <- summary(model1)
      A.Results <- Anova(model1, type = anovatype)
      RS.Results <- r.squaredGLMM(model1)
      
      plot1 <- plot(model1) #Creates diagnostic plots
      plot2 <- plot(allEffects(model1,partial.resid=TRUE)) #creates plots of model effects
      #QQPlot <- qqnorm(model1, abline = c(0,1))

      return.list <- list("Summary", S.Results,
                          "ANOVA", A.Results,
                          "R Squared", RS.Results,
                          plot1,
                          plot2 
                         # QQPlot
      )
        
      print(return.list)
}

```

##Barplots
```{r}

##Create bar plots based on single in interactive effect
BarPlots <- function(isSingleEffect, Model, PredictorVar, Ylabel="", Xlabel="") {
  
  dodge <- position_dodge(width=0.9) #For error bar positioning

#For bar graphs with a single effect (ie. no interactive effect)
  if (isSingleEffect){
    
data1 <- as.data.frame (effect(PredictorVar, Model)) #pulls data provided to function into dataframe for east of putting into model

#standard ggplot
dataplot <- (ggplot(data1, aes(x = data1[[1]], 
                                       y = data1[[2]]))
                        + theme_bw() 
                        + theme(
                                  panel.grid.major=element_blank(), #no grid lines
                                  panel.grid.minor=element_blank(), #no grid lines
                                  axis.line = element_line(colour = "black"), 
                                  panel.border=element_rect(colour = "black", fill=NA), #Add border to entire figure
                                  #set text sizes for all axis title + text
                                  axis.text.x = element_text(size = 12), 
                                  axis.title.x = element_text(size = 12),
                                  axis.title.y = element_text(size = 12),
                                  axis.text.y = element_text(size = 12),
                                  )
             #Error Bars as 95% confidence interval
                        + geom_bar(stat='identity', position=position_dodge(width=0.9))
                        + geom_errorbar(aes(ymax = data1[[2]] + (1.96*data1[[3]]), 
                                            ymin=data1[[2]] - (1.96*data1[[3]])), 
                                            position = dodge, width=0.25
                                        )
             #Axis labels as provided in function
                            + ylab(Ylabel)
                            + xlab(Xlabel)
                            + scale_fill_grey()
                              )
return(dataplot)

  }
  #For bar graphs with multiple predictors (ie. interactive effect)
  else {
    
data2 <- as.data.frame(effect(PredictorVar, Model)) #pulls data provided to function into dataframe for east of putting into model

#standard ggplot
dataplot2 <- (ggplot(data2, aes(x = data2[[2]], 
                                       y = data2[[3]], 
                                       fill = data2[[1]]))
                        + theme_bw() 
                        +  theme(
                                  panel.grid.major=element_blank(), #no grid lines
                                  panel.grid.minor=element_blank(), #no grid lines
                                  axis.line = element_line(colour = "black"),
                                  panel.border=element_rect(colour = "black", fill=NA),
                                  #set text sizes for all axis titles + text 
                                  axis.text.x = element_text(size = 12), 
                                  axis.title.x = element_text(size = 12),
                                  axis.title.y = element_text(size = 12),
                                  axis.text.y = element_text(size = 12),
                                  #Legend settings, puts in upper right corner
                                  legend.position = c(.9,.8), 
                                  legend.box.background = element_rect(colour = "black", size = 1), 
                                  legend.text = element_text(size = 12), 
                                  legend.title = element_text(size = 12), 
                                  legend.key.size = unit(.5,"line")
                                  )
              #Error bars
                        + geom_bar(stat='identity', position=position_dodge(width=0.9))
                        + geom_errorbar(aes(ymax = data2[[3]] + (1.96*data2[[4]]), 
                                            ymin=data2[[3]] - (1.96*data2[[4]])), 
                                            position = dodge, width=0.25
                                        )
              #Axis labels as provided in function
                        + labs(fill = data2[[1]]) 
                        + ylab(Ylabel)
                        + xlab(Xlabel) 
                        + scale_fill_grey() 
                              )

return(dataplot2)
}

}
```
##EnvSummary
```{r}
##Summarise Environmental data
EnvironmentalSummary <- function(EnvFile, Locations, StartTime = 0, EndTime = 24, FileName = "File.csv"){
  #Transform time data into single time format
  EnvFile = EnvFile %>% mutate(NewTime = as.POSIXct(Time, format='%H:%M:%S'),
                                NewTime2 = as.numeric(paste0(format(NewTime, '%H'), '.', format(NewTime, '%M'))))
  #Transform date to date class variable
  EnvFile$Date = as.Date(EnvFile$Date, format = "%d/%m/%y")
  
  #Filter data between start and end time provided in function, then group entries based on Date
  EnvFile = EnvFile %>% filter(between(NewTime2, StartTime, EndTime)) %>%  dplyr::group_by(Date) %>% 
  #Summarise data for each location provided by below functions
  summarise_at(vars(unlist(Locations)), 
               funs(mean, 
                    min, 
                    max, 
                    quantile(., probs = 0.25),
                    quantile(., probs = 0.75),
                    IQR
                    )
               )
 #Create list to add results of for loop into
 DataList <- list()
 
 #for each location provided in function "Locations" list create new data frame with date, mean, quantile 1 & 3, and create new column with location ID
 for(Location in Locations){
  Loc_mean <- (paste(Location, 'mean', sep='_'))
  Loc_q1 <- (paste(Location, "quantile..4", sep= '_'))
  Loc_q3 <- (paste(Location, "quantile..5", sep= '_'))
  InterQR <- (paste(Location, "IQR", sep = '_'))
  Loc <- (paste(Location))
  LocationName <- (paste(Location, "NewFile", sep = "_" ))
  
  LocationName = tibble(Date = EnvFile$Date, 
                        Mean = EnvFile[[Loc_mean]], 
                        Quantile1 = EnvFile[[Loc_q1]],
                        Quantile3 = EnvFile[[Loc_q3]],
                        InterQR = EnvFile[[InterQR]]
                        ) %>%
    add_column(Location = as.character(Loc), .after = "Date")
  
  #Add new data frame(tibble) into previously created DataList
  DataList[[Location]] <- LocationName
 }
 
 #join together rows in DataList of same name creating single dataframe with each location included 
  EnvSummForStats = bind_rows(list(DataList[[1]], DataList[[2]]))
 
  assign("EnvDataStats", EnvSummForStats, envir = .GlobalEnv) #add Dataframe with mean, Q1 & Q3 for each location provided into global environment 
  assign("EnvDataSummary", EnvFile, envir = .GlobalEnv) #Add summary statistics data frame (separate for each location) into global environment
  
  write.csv(EnvFile, file = FileName) #if file name provided in function write Summary statistic into .csv file
}
```

##DW Plots
```{r}

##Dot and whisker plots
DWPlots <- function(Model, Predictors, YLimit = c(0,100), YLabel = "", XLabel = ""){
 ( plot_model(Model, 
             terms = Predictors, 
             type = "eff",
             ci.lvl = 0.95, 
             title = "", 
             dodge = 1, 
             axis.textsize = 12)
  + ylim(YLimit)
  + labs(y= YLabel, x= XLabel)
 )
}

```



#Data input
```{r}

RTData <- (na.omit(read.csv('Data RT.csv', header=TRUE,))) #Growth measurement data
InitialData <- (na.omit(read.csv('Initial Data.csv', header=TRUE,))) #Initial volume data
Mortality <- (na.omit(read.csv('MortalityData.csv', header=TRUE,))) #Mortality data
FieldPAR <- (na.omit(read.csv('PARHOURLY.csv', header=TRUE,))) #Hourly PAR data
FieldTemp <- (na.omit(read.csv('TEMPHOURLY.csv', header=TRUE,)))#Hourly tempdata


#Change Origin and Transplant to Factor and relevel 
RTData$Origin <- as.factor(RTData$Origin)
RTData$Transplant <- as.factor(RTData$Transplant)
#RTData$Transplant <- relevel(RTData$Transplant, ref = "Shallow")
str(RTData)

#Set global theme for plot_model  
set_theme(base = theme_bw(), 
          axis.linecolor = "black",
          panel.gridcol = "white"
)


```

#Initial Bulk Volume
```{r}

#Model
InitialVol <- lm(Bulk.Volume ~ Initial.Location, data=InitialData)

Model.Diagnostics(InitialVol, 2)
#Initial Volume differed significantly between origin location, it was higher on the reef slope

```
##Plot
```{r}
set_theme(base = theme_bw(), 
          axis.linecolor = "black",
          panel.gridcol = "white"
          
)

volplotplot <-  (plot_model(InitialVol, 
                            type = "pred", 
                            terms = c("Initial.Location[Flat,Slope]"), 
                            show.ci = TRUE,
                            ci.lvl = 0.95,
                            title=" ",
                            dodge = 1,
                            axis.lim= c(2,6),
                            dot.size = 2,
                            line.size = 1,
                            show.data = TRUE,
                            axis.textsize = 12
) 
+ labs(y="Initial Bulk Volume (ml)",x="Origin")
+ geom_point(colour = "red")

)

volplotplot

```




#Mortality 
##Stats
```{r}
Mortglmer <- glmer(StatusE ~ Origin*Transplant +(1|Rock), data=Mortality, family='binomial') #Laplacian approximation
Mortglmer1 <- glmer(StatusE ~ Origin+Transplant +(1|Rock), data=Mortality, family='binomial') #Laplacian approximation
Mortglmer2 <- glmer(StatusE ~ Origin +(1|Rock), data=Mortality, family='binomial') #Laplacian approximation


AIC(Mortglmer,Mortglmer1,Mortglmer2) #Chek whether differences are significant 

Model.Diagnostics(Mortglmer2)
odds_to_rr(Mortglmer1) #RR (Rish Ratio is 0.86 - 0.86-1 = .14 so 14% less likely )

dat.sim<- simulateResiduals(Mortglmer2)
plot(dat.sim)
#want p values to be greater than 0.05)


```
##Plot
```{r}
set_theme(
  base=theme_classic(),
  axis.title.size=1,
  axis.textsize = 1#,
  #legend.size =.7,
  #legend.title.size = 3,
  #geom.label.size = 5
)

set_theme(base = theme_bw(),
          axis.linecolor = "black",
          panel.gridcol = "white"
          
)


MortPlot <- (plot_model(Mortglmer2, 
                        type = "re", 
                        terms = c("Origin[Flat,Slope]"), 
                        title=" ",
                        dodge = 1,
                        axis.textsize = 12
) 
+ labs(y="Probability of Mortality",x="Origin")

)

MortPlot


BarPlots(isSingleEffect = TRUE, Model = Mortglmer2, PredictorVar = "Origin")

summary(Mortglmer2)
summary(Mortality)

model1 <- lm(data = Mortality, StatusE ~ Origin * Transplant)

stepAIC(model1)

summary(model1)



```



#Environmental Data
##PAR Stats
```{r}
EnvironmentalSummary(FieldPAR, Locations = list("RSPAR2h", "RFPAR2H"), StartTime = 6, EndTime = 18)

PARStatsData <- as.data.frame(EnvDataStats)
PARStatsData
PARStatsData$Date <- as.factor(PARStatsData$Date)
PARStatsData$Date <- as.numeric(PARStatsData$Date)

PARFigureData <- EnvDataSummary


MeanPARLM <- lm(Mean ~ Date * Location, data = PARStatsData, na.action = "na.omit") #Best Model

stepAIC(MeanPARLM)

Model.Diagnostics(MeanPARLM, anovatype = 3)



VarPARLM <- lm(InterQR ~ Date * Location, data = PARStatsData, na.action = "na.omit")

stepAIC(VarPARLM)

Model.Diagnostics(VarPARLM, anovatype = 3)



```

##PAR Figure
```{r}
PARPlot <- (ggplot(PARFigureData, aes(x=as.Date(Date), group =1)) 
            #+ geom_point(aes(y=RS.PAR.D.mean), color = "blue") 
            + geom_line(aes(y=RSPAR2h_mean), color = "blue", group =1, size=1) 
            + geom_ribbon(aes(ymin = RSPAR2h_quantile..4,
                              ymax = RSPAR2h_quantile..5),
                          fill = "royal blue 2", alpha = 0.3, na.rm=TRUE)
            
            # + geom_point(aes(y=RF.PAR.D.mean), color = "red") 
            + geom_line(aes(y=RFPAR2H_mean), color = "red", group =1, size =1) 
            + geom_ribbon(aes(ymin = RFPAR2H_quantile..4,
                              ymax = RFPAR2H_quantile..5),
                          fill = "tomato 2", alpha = 0.3, na.rm=TRUE)
            
            + ylab(expression("PAR (μmol m"^"−2"*"s"^"−1"*")")) 
            + xlab("Date") 
            + theme_bw() 
            + theme(
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              axis.line = element_line(colour = "black"), 
              axis.text.x = element_text(angle = 90, hjust = 1, size =10),
              axis.text.y = element_text(size=12),
              axis.title.y = element_text(size = 12),
              axis.title.x = element_text(size = 12)
            ) 
            + scale_x_date(date_breaks = "6 days", 
                           limits = as.Date(c('2017-10-06','2018-02-19')))
            + scale_y_continuous(breaks=round(seq(min(0), max(1750), by = 250)), 
                                 limits = c(0,1750))
            
)

PARPlot
```

##Temp Stats
```{r echo=FALSE}
EnvironmentalSummary(FieldTemp, Locations = list("RSTempH", "RFTempH"), StartTime = 0, EndTime = 24)

TempStatsData <- as.data.frame(EnvDataStats)
TempStatsData$Date <- as.factor(TempStatsData$Date)
TempStatsData$Date <- as.numeric(TempStatsData$Date)

TempFigureData <- EnvDataSummary


MeanTempLM <- lm(Mean ~ Date * Location, data = TempStatsData, na.action = "na.omit") 
BestMeanTempLM <- lm(Mean ~ Date + Location, data = TempStatsData, na.action = "na.omit") #Best Model

stepAIC(MeanTempLM)

Model.Diagnostics(BestMeanTempLM, anovatype = 2)



VarTempLM <- lm(InterQR ~ Date * Location, data = TempStatsData, na.action = "na.omit")

stepAIC(VarTempLM)

Model.Diagnostics(VarTempLM, anovatype = 3)

plot_model(VarTempLM, 
           type = "eff")


```

##Temp Figure
```{r}
DHWValues <- (na.omit(read.csv('DHW Values.csv', header=TRUE,)))


TempFigureData %>% add_column(DHWValues$DHW)

```



```{r}
TempPlot <- (ggplot(TempFigureData, aes(x=as.Date(Date), group =1)) 
             
             #DHW data 
             + geom_line(aes(y=DHWValues$DHW +20), colour = "grey20", group = 1, size = 1, alpha = 0.3)
             + geom_ribbon(aes(ymax = DHWValues$DHW + 20, 
                               ymin = 20), 
                           fill = "grey20", alpha = 0.3)
             #Reef flat values
             # + geom_point(aes(y=RF.TEMP.mean), color = "red") 
             + geom_line(aes(y=RFTempH_mean), color = "red", group =1, size =1) 
             + geom_ribbon(aes(ymin = RFTempH_quantile..4,
                               ymax = RFTempH_quantile..5),
                           fill = "tomato 2", alpha = 0.5, na.rm=TRUE)
             #Reef slope values
             #+ geom_point(aes(y=RS.TEMP.mean), color = "blue") 
             + geom_line(aes(y=RSTempH_mean), color = "blue", group =1, size=1) 
             + geom_ribbon(aes(ymin = RSTempH_quantile..4,
                               ymax = RSTempH_quantile..5),
                           fill = "royal blue 2", alpha = 0.5, na.rm=TRUE)
             
             
             + ylab(expression("Temperature (°C)")) 
             + xlab("Date") 
             + theme_bw() 
             + theme(
               panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black"), 
               axis.text.y = element_text(size=12),
               axis.title.y = element_text(size = 12, vjust = 2),
               axis.text.x = element_text(size = 12, angle = 90, hjust = 1),
               axis.title.x = element_text(size = 12)
             ) 
             + scale_x_date(date_breaks = "6 days", 
                            limits = as.Date(c('2017-10-06','2018-02-19')))
             #  + scale_y_continuous(breaks=round(seq(min(20), max(32), by = 2)),
             # limits = c(20,32))
             + scale_y_continuous(DHWValues$DHW, name = "Temperature (°C)", 
                                  sec.axis = sec_axis(~.-20, name = "DHW (°C Week)"),
                                  breaks=round(seq(min(20), max(32), by = 2)),
                                  limits = c(20,32))
             + geom_hline(yintercept = 28.3, linetype = "dashed", size = 1)
             
)


TempPlot
```




#Bulk Volume ~ Origin*Transplant
##Stats
```{r}
#The Linear mixed effects model includes "rock" as a random factor.
VolLME <- lme(Pct.Volume ~ Origin  * Transplant, 
              random=~1|Rock, data=RTData, method="ML", na.action = na.exclude)

#Model selection stepAIC
stepAIC(VolLME)

#take forumula for best fit model, and rerun lme, with these predictors and relationships, but this time use method = "REML"
#eg. IF only location remains, 
VolLME.auto <- lme(Pct.Volume ~ Origin + Transplant, 
                   random=~1|Rock, data=RTData, method="REML", na.action = na.exclude)
#reason is that cannot use stepAIC using REML. See BIOL3023 manual

#model diagnostics
Model.Diagnostics(VolLME.auto, 2)
qqnorm(VolLME.auto, abline = c(0,1))

#fixed effects account for 37% of variance, random effects an additional 7%
#Corals moved to the shallows increase volume more than those moved to the deep, but reef-flat corals increase volume less than reef slope corals

```

##Plots
```{r}
#Origin Bar Plot
VolumeOriginPlot <- BarPlots(TRUE, VolLME.auto, "Origin",
                             Ylabel = "Bulk Volume (ml)",
                             Xlabel = "Origin Location")
#Transplant Bar Plot
VolumeTransplantPlot <- BarPlots(TRUE, VolLME.auto, "Transplant", 
                                 Ylabel = "%\U0394 Bulk Volume (ml)",
                                 Xlabel = "Transplant Locations")



#Transplant dot and whisker plot
VolumeTransplantPlot2 <- DWPlots(VolLME.auto, "Transplant", 
                                 YLimit = c(20,100), 
                                 YLabel = "%\U0394 Bulk Volume (ml/ml)", 
                                 XLabel = "Transplant Locations"
)

#Origin dot and whisker plot
VolumeOriginPlot2 <- DWPlots(VolLME.auto, "Origin", 
                             YLimit = c(20,100), 
                             YLabel = "%\U0394 Bulk Volume (ml/ml)", 
                             XLabel = "Origin Locations"
)

print(list(VolumeOriginPlot, VolumeOriginPlot2, 
           VolumeTransplantPlot, VolumeTransplantPlot2))




```


#Surface Area 
##Stats
```{r}
SALME <- lme(Pct.SA ~ Origin * Transplant + Pct.Volume, 
             random=~1|Rock, data=RTData, method="ML", na.action = na.exclude)

#Model selection stepAIC
stepAIC(SALME)

#take forumula for best fit model, and rerun lme, with these predictors and relationships, but this time use method = "REML"
#eg. IF only location remains, 
SALME.auto <- lme(Pct.SA ~ Origin + Transplant + Pct.Volume, 
                  random=~1|Rock, weights = varIdent(form = ~1 | Origin * Transplant), data=RTData, method="REML", na.action = na.exclude)
#reason is that cannot use stepAIC using REML. See BIOL3023 manual

#fixed effects account for 77% of variance, random effects an additional 0%

#model diagnostics
Model.Diagnostics(SALME.auto, 2)
qqnorm(SALME.auto, abline = c(0,1))

# increase tissue for fixed increase in  branch volume  (living SA to Volume) is less for corals placed on reef slope, but greater for corals originating from reef slope

```

##Plots
```{r}

#Origin bar plot
SAOriginPlot <- BarPlots(TRUE, SALME.auto, "Origin", 
                         Ylabel = "% Living Surface Area", 
                         Xlabel = "Origin Location"
)


#Origin dot and whisker plot
SAOriginPlot2 <- DWPlots(SALME.auto, "Origin", YLimit = c(40,80), 
                         YLabel = expression("%\U0394 LSA (cm"^"2"*"/cm"^"2"*")"), 
                         XLabel = "Origin Locations"
)

#Transplant bar plot
SATransplantPlot <- BarPlots(TRUE, SALME.auto, "Transplant", 
                             Ylabel = expression("%\U0394 LSA (cm"^"2"*"/cm"^"2"*")"), 
                             Xlabel = "Transplant Locations"
)

#Transplant dot and whisker plot
SATransplantPlot2 <- DWPlots(SALME.auto, "Transplant", YLimit = c(40,80), 
                             YLabel = expression(bold("%\U0394 LSA (cm"^"2"*"/cm"^"2"*")")), 
                             XLabel = "Transplant Locations"
)

print(list(SAOriginPlot, SAOriginPlot2, 
           SATransplantPlot, SATransplantPlot2))







```
#Buoyant Weight
##Stats
```{r}

BWvolLME <- lme(Pct.BW ~ Origin * Transplant+ Pct.Volume + Pct.SA, 
                random=~1|Rock, data=RTData, method="ML", na.action = na.exclude)


#Model selection stepAIC
stepAIC(BWvolLME)

#take forumula for best fit model, and rerun lme, with these predictors and relationships, but this time use method = "REML"
#eg. IF only location remains, 
CalvolLME.auto2 <- lme(Pct.BW ~ Origin * Transplant  + Pct.SA ,
                       random=~1|Rock, weights = varIdent(form = ~1 | Origin * Transplant),data=RTData, method="REML", na.action = na.exclude)
#reason is that cannot use stepAIC using REML. See BIOL3023 manual
#Have added weights as improve assumption fit

vif(CalvolLME.auto2) # OK if <5 as standard for collinearity, otherwise drop SA from model, get same effect

#model diagnostics
Model.Diagnostics(BWvolLME, 3)
qqnorm(CalvolLME.auto2, abline = c(0,1))

#Density increase of branch  (Calcification to Volume for same change in living tissue) is greater for reef flat corals returned to reef flat.

```
##Plots
```{r}

set_theme(base = theme_bw(), 
          axis.linecolor = "black",
          panel.gridcol = "white",
          legend.inside = TRUE,
          legend.pos = "top right",
          legend.just = c(1.2,1.1),
          legend.item.backcol = "white"
)

PCTcalplot <- (plot_model(CalvolLME.auto2, 
                          type = "eff",
                          terms = c("Origin", "Transplant"),
                          ci.lvl = 0.95, 
                          title = "", 
                          dodge = c(.5), 
                          axis.textsize = 12, 
                          show.legend = TRUE,
                          colors = c("blue","red")
)
               + ylim(80,140)
               + labs(y="%\U0394 Buoyant Weight (g/g)",x="")
               + theme(
                 axis.text.x = element_blank(),
                 axis.title.x = element_blank()
               ))

PCTcalplot

```


#Protein
##Stats
```{r}
#The Linear mixed effects model includes "rock" as a random factor.
ProteinLME <- lme(Pct.Protein ~ Origin * Transplant + Pct.BW + Pct.SA + Pct.Volume, 
                  random=~1|Rock, data=RTData, method="ML", na.action = na.exclude)

summary(ProteinLME) 

#Model selection stepAIC
ProteinLME.auto <-stepAIC(ProteinLME)

# summaries:
summary(ProteinLME.auto)
#take forumula for best fit model, and rerun lme, with these predictors and relationships, but this time use method = "REML"
#eg. IF only location remains, 
ProteinLME.auto2 <- lme(Pct.Protein ~ Origin * Transplant + Pct.BW +Pct.SA + Pct.Volume,
                        random=~1|Rock, weights = varIdent(form = ~1 | Origin * Transplant), data=RTData, method="REML", na.action = na.exclude)
#reason is that cannot use stepAIC using REML. See BIOL3023 manual

#model fit:
r.squaredGLMM(ProteinLME.auto2)  #fixed effects account for 47% of variance, random effects an additional 0%



plot(ProteinLME.auto2)

qqnorm(ProteinLME.auto2, abline = c(0,1))
summary(ProteinLME.auto2)
Anova(ProteinLME.auto2, type = 3)
plot(allEffects(ProteinLME.auto2,partial.resid=TRUE))
lsmeans(ProteinLME.auto2,pairwise ~ Origin*Transplant,adjust="tukey") #Tukey post hoc test to determine levels of variance.

emm.s <- emmeans(ProteinLME.auto2, c("Origin","Transplant"))
pairs(emm.s)

```



##Plots
```{r}

# set_theme(base = theme_bw(),
# axis.linecolor = "black",
# panel.gridcol = "white"
# )

PCTprotplot <- (plot_model(ProteinLME.auto2, 
                           type = "eff",
                           terms = c("Origin", "Transplant"),
                           ci.lvl = 0.95, 
                           title = "", 
                           dodge = .5, 
                           axis.textsize = 12,
                           show.legend = FALSE,
                          colors = c("blue","red"))
                + ylim(40,120)
                + labs(y= expression(bold("%\U0394 Total Protein (mg/mg)"),x="Origin Locations"))
)
PCTprotplot


```
#Arrange and save figure 1
```{r}

TempPlot <- (TempPlot
             + theme(
               axis.text.x = element_blank(),
               axis.title.x = element_blank()
               
             )
)

top1 <- ggpubr::ggarrange(TempPlot, PARPlot,
                          align = "v",
                          ncol = 1,
                          heights = c(15,20),
                          labels = c("c","d"), 
                          label.x = .11,
                          label.y = 0.96
)

bottom1 <- ggpubr::ggarrange(MortPlot, volplotplot,
                             align = "h",
                             ncol=2,
                             labels = c(" a","b"), 
                             label.x = c(.19, .12),
                             label.y = 0.9
)
bottom1


Figure1 <- cowplot::plot_grid(bottom1, top1,
                              align = "vh",
                              rel_heights = c(1,1.5),
                              ncol = 1)

Figure1




ggsave("Figure1.png", 
       plot = Figure1, 
       path = "Figures", 
       scale = 1, 
       width = 8, 
       height = 8, 
       units = c("in", "cm", "mm"),
       dpi = 1000)


```


#Arrange and Save Growth Panel graph 
```{r}
VolumeTransplantPlot2 <- (VolumeTransplantPlot2
                          + theme(
                            axis.title.y = element_text(face = "bold"),
                            axis.text.x = element_blank(), 
                            axis.title.x = element_blank(),
                            panel.background = element_rect(fill="transparent"),
                            plot.background = element_rect(fill = "transparent", colour = NA),
                            panel.grid.major = element_blank(), # get rid of major grid
                            panel.grid.minor = element_blank()
                          )
)

VolumeOriginPlot2 <- (VolumeOriginPlot2
                      + theme(
                        axis.text.x = element_blank(), 
                        axis.title = element_blank(),
                            panel.background = element_rect(fill="transparent"),
                            plot.background = element_rect(fill = "transparent", colour = NA),
                            panel.grid.major = element_blank(), # get rid of major grid
                            panel.grid.minor = element_blank()
                      )
)

SAOriginPlot2 <- (SAOriginPlot2
                      + theme(
                        axis.title.x = element_text(face = "bold"),
                        axis.title.y = element_blank(),
                            panel.background = element_rect(fill="transparent"),
                            plot.background = element_rect(fill = "transparent", colour = NA ),
                            panel.grid.major = element_blank(), # get rid of major grid
                            panel.grid.minor = element_blank()
                      )
)

SATransplantPlot2 <- (SATransplantPlot2
                      + theme(
                            axis.title.y = element_text(face = "bold"),
                            axis.title.x = element_text(face = "bold"),
                            panel.background = element_rect(fill="transparent"),
                            plot.background = element_rect(fill = "transparent", colour = NA),
                            panel.grid.major = element_blank(), # get rid of major grid
                            panel.grid.minor = element_blank()
                      )
)

PCTcalplot <- (PCTcalplot
                  + theme(
                          axis.title.y = element_text(face = "bold"),
                          panel.background = element_rect(fill = "transparent"), # bg of the panel
                          plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
                          panel.grid.major = element_blank(), # get rid of major grid
                          panel.grid.minor = element_blank(), # get rid of minor grid
                          legend.background = element_rect(fill = "transparent", colour = NA), # get rid of legend bg
                          legend.box.background = element_rect(fill = "transparent", colour = NA), 
                          legend.key = element_rect(fill = "transparent", colour = NA)
                  )

)

PCTprotplot <- (PCTprotplot
                + theme(
                  axis.title.y = element_text(face = "bold"),
                  axis.title.x = element_text(face = "bold"),
                            panel.background = element_rect(fill="transparent"),
                            plot.background = element_rect(fill = "transparent", colour = NA),
                            panel.grid.major = element_blank(), # get rid of major grid
                            panel.grid.minor = element_blank()
                      )
)

                      


top2 <- ggpubr::ggarrange(VolumeTransplantPlot2, VolumeOriginPlot2, 
                          SATransplantPlot2, SAOriginPlot2,
                          align = "hv",
                          labels = c("a","b"," c","d"),
                          label.x = c(.3,.3,.28,.3),
                          label.y = .86
)


bottom2 <- ggpubr::ggarrange(PCTcalplot, PCTprotplot, 
                             heights = c(10,11.5),
                             align = "v",
                             ncol=1,
                             labels = c("e"," f"),
                             label.x = .15,
                             label.y = .85
)

figure2 <- cowplot::plot_grid(top2, bottom2, 
                                align = "hv",
                                ncol = 1, 
                                greedy = FALSE)

figure2


ggsave("Figure2.png",
       plot = figure2,
       path = "Figures",
       scale = 1,
       width = 5,
       height = 12,
       units = c("in", "cm", "mm"),
       dpi = 1000, 
       bg = "transparent")

```